---
title: "Lab 10: Filtering Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
theme_set(theme_minimal())
```

## Pick a dataset

Today you will again have a choice of three datasets to work with. You can
choose either to investigate the daily team ELO ratings in the NBA:

```{r}
nba <- read_csv("https://statsmaths.github.io/stat_data/nba_elo_daily.csv")
```

```{r}
nba_21_century <- filter(nba, season >= 2000)
nba_21_century

ggplot(nba_21_century, aes(season, elo)) +
  geom_point(aes(color = team))
```

```{r}
nba_lakers_21 <- filter(nba, team == "LAL", season >= 2000)
nba_lakers_21
nba_lakers_21_higer <- filter(nba, team == "LAL", season >= 2000, elo > 1700)

ggplot(nba_lakers_21, aes(date, elo)) +
  geom_point() +
  geom_point(data = nba_lakers_21_higer, color = "red")


```

```{r}
nba_super_team <- filter(nba, elo > 1700)
nba_super_team
ggplot(nba_super_team, aes(team, date)) +
  geom_point(aes(color=team))

```


## Instructions

This lab is similar to the last. You'll pick a dataset and then generate an
analysis of that data using data visualizations. The difference here is that
instead of a single plot, I would like you to construct 3-4 plots that
together tell a linear story. Each plot should be separated by a sentence or
two describing what the viewer should take away from the plot. Try to keep the
plots similar in some way; perhaps each simply highlights a different subset
of the data but has the same underlying layers.

Notice that each of the datasets for today are larger than you will probably
be able to use. You will need to filter the data to a particular team, year,
subset of curse words. You may further filter the data to highlight an even
smaller subset of the data within each plot.

## Analysis










