---
title: "nba three pointer precentage and winning precentage "
author: "Yiming Fan"
output: html_notebook
---





```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(plotly)
theme_set(theme_minimal())
```

Is there a correlation betwween 3-point shots ability and winning precentage?

## Overview

PUT AN INTRODUCTORY PARAGRAPH HERE GIVING IN 4-5 SENTENCES
THE BASIC IDEA OF YOUR DATA COLLECTION TASK AND HOW YOU WENT
ABOUT IT. INCLUDE THE OVERARCHING RESEARCH QUESTION DESCRIBING
WHY YOU WERE INTERESTED IN THIS DATA

Since I was a child, I have watched the NBA. Nowadys, the basketball league is in the middle of a revolution that more and more teams are willing to take 3-point shots. In my project, I try to explore if taking 3-point shots can really maximize offensive efficiency by improving each team's winning precentage. In other words, I want to know if there is a correlation between 3-point shots ability and winning precentage.

## Data dictionary


- **season**: a numeric variable describing which year of stats we are using.
- **team_name**:a catagorical variable describing each team's name. 
options from:
Houston
Brooklyn
Cleveland
Toronto
Dallas
Boston
Chicago
Miami
Denver
Atlanta
Golden State
Oklahoma City
New Orleans
Orlando
Utah
Philadelphia
Detroit
LA Clippers
LA Lakers
Phoenix
Portland
Memphis
Charlotte
Washington
Indiana
San Antonio
Milwaukee
Sacramento
Minnesota
New York
- **number_of_3pointer_attempts**: a numeric variable describing 3 point shots attempted. 
- **number_of_made**:a numeric variable describing successful 3 pointer attempts.
- **3pointer_percentage**:a numeric variable dividing made 3 pointer by total 3 pointer attempted.
- **winning_precentage**:a numeric variable dividing game won by total game played. 





```{r}
nba <- read_csv("https://raw.githubusercontent.com/URstats/stat209-s18-fanyiming33/master/nba.csv") 
nba
```
intext citation (ESPN)


## Exploratory Analysis
As we have seen in the nba, those teams who are good at shooting threes usually have a better precentage of winning the game. For the recent five years, we can find that teams are more liely to attempt 3 pointers. Our task in this project is to find the real reason why most of the teams are attempting 3 pointers knowadays.
We will dig into the relationship between 3 pointers and winning record on my first graph. Then, we will analyze the winning precentage and three pointer attempted of Houston which is the best 3 pointer team in the league.


```{r, echo=FALSE}
nba_2017 <- filter(nba, season == 2017)
nba_2017
ggplot(nba_2017, aes(threepointer_precentage, winning_precentage)) +
  geom_point(aes(color=team_name), size=5) +
  ggtitle("the effect of threepointer precentage on winning precentage") +
  annotate("text", x = 0.386, y = 0.759, label = "the best 3pointer team" ) +
  annotate("text", x = 0.366, y = 0.772, label = "the team winning the most") +
  annotate("text", x = 0.342, y = 0.305, label = "worst team and 3pointer team")

ggplotly()
```

```{r}
rocket <- filter(nba, team_name == "Houston")
rocket
ggplot(rocket, aes(season, winning_precentage)) +
  geom_point(aes(size=number_of_threepointer_attempts), color="red") +
  geom_line()

```

## Conclusion

From the first graph, we can clearly see a relationship between each team's winning precentage and threepointer's precentage. As more 3 shots are made, a team has a greater chance of winning. In the second graph, we can conclude from Houston that the increasing amount of three pointer increse the possibility of winning. 

## References

"NBA Statistics and League Leaders." NBA Stats,       sports.espn.go.com/nba/statistics?season=2009&seasontype=2.


